Powdery mildew yield loss mitigation strategies in mungbean, a meta-analysis.

### 2019-11-05 04_Prep_data.Rmd
Updated to incorportate data wrangling from the means data set to a slim data with clustered spray categories

@@

# dat1 <- PM_MB_means[PM_MB_means$trial_ref == "AM1303", ]
# 
# controlY <-
#    mean(dat1[dat1$fungicide == "control", "grain_yield.t.ha."])
# 
# dat1$yield_gain <- dat1$grain_yield.t.ha. - controlY
# dat1$prop_YG <- dat1$yield_gain / controlY
# dat1[dat1$fungicide == "control", c("yield_gain", "prop_YG")] <- NA
# PM_MB_means[PM_MB_means$trial_ref == "AM1303", ] <- dat1

write.csv(PM_MB_means,
          "data/1902 PM_MB_means&Ygains.csv")
```

@@

slim_PM_dat[slim_PM_dat$fungicide_ai == "control", c("fungicide_timing_1", "fungicide_timing_2", "fungicide_timing_3",
                                                     "cluster_1", "cluster_2","cluster_3")] <- "control"

>>>>

## Standardising sample variance

The type of variance needs to be standardised between trials and treatments for the meta-analysis.

```{r variance_types}
# PM_MB_means$Y_error_type <- forcats::fct_explicit_na(PM_MB_means$Y_error_type, na_level = NA)
# Trials using tebuconazole
PM_MB_means %>%
   filter(fungicide_ai == "tebuconazole",
         row_spacing == 0.75) %>%
   group_by(trial_ref, location, Y_error_type) %>%
   summarise(tebuconazole_trts = length(Y_error_type))

# Trials using propiconazole
PM_MB_means %>%
   filter(fungicide_ai == "propiconazole",
          row_spacing == 0.75) %>%
   group_by(trial_ref, location, Y_error_type) %>%
   summarise(propiconazole_trts = length(Y_error_type))
```

The experiment summaries report either standard deviation, least square differences or no error term (`NA`). Below we will convert LSD and standard deviation to sample variance

### Converting least-squares to sample variance

```{r all_LSD_trials}
Trials <- PM_MB_means %>%
   filter(
          Y_error_type == "lsd (P=0.05)") %>%
   select(trial_ref) %>%
   distinct()
Trials
```

The trial, `mung1112/02` uses Least Square Differences (LSD) to describe the variation within the experiment.
The following code attempts to convert this to sampling variance as per the method in [Nugugi et.al (2011)](https://apsjournals.apsnet.org/doi/10.1094/PHYTO-08-10-0221).
Ideas for using a T-critical value of 1.697 came from reading a [statisics-how-to website](https://www.statisticshowto.datasciencecentral.com/how-to-calculate-the-least-significant-difference-lsd/).
A [table of T-critical values](https://www.statisticshowto.datasciencecentral.com/tables/t-distribution-table/) was consulted where 30 degrees of freedom within experiment was used to find the 'T-crit' value of 1.697.

![Calculating sampling variance from mean squared error Paul et al. (2008)](Paul_etal_2008.PNG)

```{r LSD_2_Sampling Variance}
# Formula and modifications based off information found at https://www.statisticshowto.datasciencecentral.com/how-to-calculate-the-least-significant-difference-lsd/

for (Trial2012 in Trials$trial_ref) {
   Tcrit <-
      1.697  # 0.05 and DFw = 30 # T critical value for which there is significant difference between two groups; This value is the same for both trials "mung1112/01" and "mung1112/02"
   DFw <-
      sum(PM_MB_means[PM_MB_means$trial_ref == Trial2012, ]$replicates) -
      length(PM_MB_means[PM_MB_means$trial_ref == Trial2012, ]$replicates) # degrees of freedom within groups
   # n.A <- PM_MB_means[PM_MB_means$trial_ref == "Trial2012,]$replicates[1]
   # n.B <- PM_MB_means[PM_MB_means$trial_ref == "Trial2012,]$replicates[2]
   LSD <-
      PM_MB_means[PM_MB_means$trial_ref == Trial2012, ]$yield_error[1]

   # SSwithin <- #unknown
   # MSW <- SSwithin / DFw # mean square within (SS within / DF within)

   #  LSD = (Tcrit) * sqrt(MSW * (1/n.A + 1/n.B))
   #  (LSD/Tcrit) = sqrt(MSW * (1/n.A + 1/n.B))
   #  (LSD/Tcrit)^2 = MSW * (1/n.A + 1/n.B)
   #  (LSD/Tcrit)^2/(1/n.A + 1/n.B) = MSW
   #  MSE <- (LSD/Tcrit)^2/(1/n.A + 1/n.B) # This formula was abandon in favor for the formula in Nugugi et. at (2011)
   MSE <-
      PM_MB_means[PM_MB_means$trial_ref == Trial2012, ]$replicates[1] * (LSD /
                                                                            Tcrit) ^ 2 / 2 # method for finding variance in Nugugi et.al (2011) DOI: 10.1094/phyto-08-10-0221

   for (j in seq_along(PM_MB_means[PM_MB_means$trial_ref == Trial2012, ]$replicates)) {
      if (j == 1) {
         Sv <-
            vector(length = length(PM_MB_means[PM_MB_means$trial_ref == Trial2012, ]$replicates))
      }
      #  Sv[j]^2 = MSE/(PM_MB_means[PM_MB_means$trial_ref == Trial2012,]$replicates[j] *
      #                   PM_MB_means[PM_MB_means$trial_ref == Trial2012,]$grain_yield.t.ha[j])
      # Balance equation
      Sv[j] = sqrt(MSE / (
         PM_MB_means[PM_MB_means$trial_ref == Trial2012, ]$replicates[j] *
            as.numeric(PM_MB_means[PM_MB_means$trial_ref == Trial2012, ]$grain_yield.t.ha[j])
      ))

   }

   # and replacing the values in the dataset

   PM_MB_means[PM_MB_means$trial_ref == Trial2012, ]$yield_error <-
      Sv
   PM_MB_means[PM_MB_means$trial_ref == Trial2012, ]$Y_error_type <-
      "SamplingVariance"

}
```

### Converting standard deviation to sample variance
Now to estimate the variance from the standard deviation

```{r stdev2SampVariance}
# subset data by those which describe the Yield error with standard deviation
Trials <- PM_MB_means %>%
   filter(#fungicide_ai == "tebuconazole",
      #row_spacing.m == 0.75,
      Y_error_type == "stdev") %>%
   select(trial_ref) %>%
   distinct()

# calculate the Sampling variance from the standard deviation
# standard error of the mean equals the standard deviation divided by the square root of the number of samples
# Sv <- sem^2 <- stdev/sqrt(n)
# Sampling variance equals the square of the standard error.
# Sv <- (stdev/sqrt(n))^2

# Therefore

for (j in seq_along(PM_MB_means[PM_MB_means$trial_ref %in% Trials$trial_ref,]$yield_error)) {
   # create empty vector at first iteration with length of j
   if (j == 1) {
      Sv <-
         vector(length = length(PM_MB_means[PM_MB_means$trial_ref %in% Trials$trial_ref,]$yield_error))
   }

   # Equation to convert standard deviation to Sampling Variance
   Sv[j] = (PM_MB_means[PM_MB_means$trial_ref %in% Trials$trial_ref,]$yield_error[j]) ^
      2
}

# and replacing the values in the dataset
PM_MB_means[PM_MB_means$trial_ref %in% Trials$trial_ref,]$yield_error <-
   Sv
PM_MB_means[PM_MB_means$trial_ref %in% Trials$trial_ref,]$Y_error_type <-
   "SamplingVariance"
```

## Yield gains  
Yield gains and yield gains proportional to the control.

@@

dat1 <- PM_MB_means[PM_MB_means$trial_ref == "AM1303", ]

controlY <-
   mean(dat1[dat1$fungicide == "control", "grain_yield.t.ha."])

dat1$yield_gain <- dat1$grain_yield.t.ha. - controlY
dat1$prop_YG <- dat1$yield_gain / controlY
dat1[dat1$fungicide == "control", c("yield_gain", "prop_YG")] <- NA
PM_MB_means[PM_MB_means$trial_ref == "AM1303", ] <- dat1

write.csv(PM_MB_means,
          "data/1911_PM_MB_means&Ygains.csv")
```

@@

slim_PM_dat[slim_PM_dat$fungicide_ai == "control",
            c("fungicide_timing_1", "fungicide_timing_2", "fungicide_timing_3",
              "cluster_1", "cluster_2","cluster_3"
              )] <- "control"

write.csv(slim_PM_dat, file = "data/slim_PM_clusterdat.csv", row.names = FALSE)
